spatial_figures

### 2025-03-31 spatial_figs.Rmd
climate diff figs

```{r}
grid <- read_csv("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/grid_all_Catchment.csv",
                 skip=3) %>%
   slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(date = ymd(time)) %>%
  dplyr::select(-time) %>%
  mutate_if(is.character, as.numeric) %>% 
  mutate(site = 'grid')

catchment_mo_mean <- all %>%
  mutate(month = month(date)) %>%
  group_by(site, month) %>%
  summarise(across(precip:solRad, ~ mean(.x, na.rm = TRUE))) %>%
  dplyr::select(site, month,
                precip_mean = precip, tmean_mean = tmean, 
                solrad_mean = solRad, rh_mean = rhum)

catchment_mo_sd <- all %>%
  mutate(month = month(date)) %>%
  group_by(site, month) %>%
  summarise(across(precip:solRad, ~ sd(.x, na.rm = TRUE))) %>%
  dplyr::select(site, month,
                precip_sd = precip, tmean_sd = tmean, 
                solrad_sd = solRad, rh_sd = rhum)

p_cols <- ggplot(catchment_mo2, aes(x = month_name, y = precip_mean,
                                    color = site, fill = site)) +
  geom_col(position = 'dodge') +
  geom_errorbar(aes(ymin = 0, ymax = precip_mean + precip_sd), 
                position = position_dodge(width = 0.6), width = 0.2) +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Rename color legend items
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Rename fill legend items
  theme_bw(base_size = 20) +
  labs(y = 'Precipitation (mm)',
       x = element_blank()) +
  guides(color = guide_legend(title = NULL),
         fill = guide_legend(title = NULL)) +
    theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank()
  )

ggsave("P.png", plot = p_cols, width = 8, height = 6, dpi = 300)

# Temperature plot
t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean, 
                                      color = site, group = site)) +
  geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.2) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size=12) +
  labs(y = expression("Mean temperature " ( degree~C)),
       x = element_blank()) +
  labs(color = NULL, fill = NULL)

# Solar radiation plot
sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean, 
                                         color = site, group = site)) +
  geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.2) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size = 12) +
  labs(y = expression("Solar radiation (MJ/m"^"2" * ")"),
       x = element_blank()) +
  labs(color = NULL, fill = NULL)

# Relative humidity plot
rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean, 
                                        color = site, group = site)) +
  geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.2) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size = 12) +
  labs(y = 'RH (%)', x= element_blank()) +
  labs(color = NULL, fill = NULL)

# Print the plots
t_moplot
sol_moplot
rh_moplot

# t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean, 
#                                       color = site, group = site)) +
#   geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   theme_bw(base_size=14) +
#   labs(y = expression("Mean temperature " ( degree~C)),
#        x = element_blank())
# 
# t_moplot
# 
# sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean, 
#                                       color = site, group = site)) +
#   geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   theme_bw(base_size = 14) +
#   labs(y = expression("Solar radiation (MJ/m"^"2" * ")"),
#        x = element_blank()) 
# 
# sol_moplot
# 
# 
# rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean, 
#                                       color = site, group = site)) +
#   geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   theme_bw(base_size = 14) +
#   labs(y = 'RH (%)', x= element_blank()) 
# rh_moplot
library(ggpubr)

ggsave("./figures/quad.png", plot = quad, dpi = 600)

quad_plus <- ggarrange(flow_plot,quad,
                       nrow=2)
quad_plus

ggsave("./figures/quad_plus.png", plot = quad_plus, dpi = 600)

@@

# difference - UPDATE file sources!!

@@

statn_catch <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_model/output/4milecreek38HRUs-statnall_#1/out/Catchment.csv',
                        skip=3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'statn')

grid_catch <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_model/output/4milecreek38HRUs-gridall_#1/out/Catchment.csv',
                       skip=3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'grid')

catchment_p <- bind_rows(statn_catch, grid_catch) %>%
  dplyr::select(precip, time, source) %>%
  pivot_wider(., names_from=source, values_from=precip)

p_comp <- ggplot(catchment_p, aes(x=statn, y=grid)) + geom_point(size = 3) +
  theme_bw(base_size=25) +
  labs(x = 'Station precip (mm)',
       y = 'Grid precip (mm)') +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")

p_comp
ggsave("./figures/p_compare.png", plot = p_comp, dpi = 600)

@@

p_cols <- ggplot(catchment_mo2, aes(x = month_name, y = avg_p,
                                    color = site, fill = site)) +
  geom_col(position = 'dodge') +
  geom_errorbar(aes(ymin = 0, ymax = avg_p + sd_p), 
                position = position_dodge(width = 0.6), width = 0.2) +
  scale_color_manual(values = c("#0047AB", "orange"),
                     labels = c("Gridded", "Station")) +  # Rename color legend items
  scale_fill_manual(values = c("#0047AB", "orange"),
                    labels = c("Gridded", "Station")) +   # Rename fill legend items
  theme_bw(base_size = 20) +
  labs(y = 'Precipitation (mm)',
       x = element_blank()) +
  guides(color = guide_legend(title = NULL),
         fill = guide_legend(title = NULL)) +
    theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank()
  )

@@

ggsave("/Users/megansears/Documents/Repos/fourmile/figures/quad.png", plot = quad, dpi = 600)

>>>>

# grid <- read_csv("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/grid_all_Catchment.csv",
#                  skip=3) %>%
#    slice(3:n()) %>%
#   dplyr::select(-starts_with('@'),
#                 -1) %>%
#   mutate(date = ymd(time)) %>%
#   dplyr::select(-time) %>%
#   mutate_if(is.character, as.numeric) %>% 
#   mutate(site = 'grid')
# 
# statn <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/statn_all_Catchment.csv',
#                  skip=3) %>%
#    slice(3:n()) %>%
#   dplyr::select(-starts_with('@'),
#                 -1) %>%
#   mutate(date = ymd(time)) %>%
#   dplyr::select(-time) %>%
#   mutate_if(is.character, as.numeric) %>% 
#   mutate(site = 'statn')
# 
# all <- bind_rows(grid, statn) %>%
#   arrange(date)
# 
# catchment_mo_mean <- all %>%
#   mutate(month = month(date)) %>%
#   group_by(site, month) %>%
#   summarise(across(precip:solRad, ~ mean(.x, na.rm = TRUE))) %>%
#   dplyr::select(site, month,
#                 precip_mean = precip, tmean_mean = tmean, 
#                 solrad_mean = solRad, rh_mean = rhum)
# 
# catchment_mo_sd <- all %>%
#   mutate(month = month(date)) %>%
#   group_by(site, month) %>%
#   summarise(across(precip:solRad, ~ sd(.x, na.rm = TRUE))) %>%
#   dplyr::select(site, month,
#                 precip_sd = precip, tmean_sd = tmean, 
#                 solrad_sd = solRad, rh_sd = rhum)
# 
# 
# catchment_mo <- left_join(catchment_mo_mean, catchment_mo_sd, 
#                           by = c('month', 'site'))
# 
# catchment_mo2 <- catchment_mo %>%
#   mutate(month_name = month.abb[month]) %>%
#   mutate(month_name = factor(month_name, 
#                              levels = c('Jan', 'Feb', 'Mar',
#                                         'Apr', 'May', 'Jun',
#                                         'Jul', 'Aug', 'Sep',
#                                         'Oct', 'Nov', 'Dec')))
# 
# 
# p_cols <- catchment_mo2 %>% 
#   filter(month_name %in% c('Jun','Jul','Aug', 'Sep')) %>%
#   ggplot(., aes(x = month_name, y = precip_mean,
#                                     color = site, fill = site)) +
#   geom_col(position = 'dodge') +
#   geom_errorbar(aes(ymin = 0, ymax = precip_mean + precip_sd), 
#                 position = position_dodge(width = 0.6), width = 0.2) +
#   scale_color_manual(values = c("blue", "orange"),
#                      labels = c("Gridded", "Observed")) +  # Rename color legend items
#   scale_fill_manual(values = c("blue", "orange"),
#                     labels = c("Gridded", "Observed")) +   # Rename fill legend items
#   theme_bw(base_size = 20) +
#   labs(y = 'Precipitation (mm)',
#        x = element_blank()) +
#   guides(color = guide_legend(title = NULL),
#          fill = guide_legend(title = NULL)) +
#     theme(
#     panel.grid = element_blank(),
#     panel.background = element_rect(fill = "white"),
#     text = element_text(color = "black"),
#     axis.title = element_text(color = "black"),
#     axis.text = element_text(color = "black"),
#     strip.text = element_blank()
#   ) 
# 
# 
# p_cols
# 
# ggsave("P.png", plot = p_cols, width = 8, height = 6, dpi = 300)
# 
# # Temperature plot
# t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean, 
#                                       color = site, group = site)) +
#   geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   scale_color_manual(values = c("blue", "orange"),
#                      labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
#   scale_fill_manual(values = c("blue", "orange"),
#                     labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
#   theme_bw(base_size=20) +
#   labs(y = expression("Mean temperature " ( degree~C)),
#        x = element_blank()) +
#    guides(color = guide_legend(title = NULL),
#          fill = guide_legend(title = NULL)) +
#     theme(
#     panel.grid = element_blank(),
#     panel.background = element_rect(fill = "white"),
#     text = element_text(color = "black"),
#     axis.title = element_text(color = "black"),
#     axis.text = element_text(color = "black"),
#     strip.text = element_blank()
#   ) 
# 
# # Solar radiation plot
# sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean, 
#                                          color = site, group = site)) +
#   geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   scale_color_manual(values = c("blue", "orange"),
#                      labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
#   scale_fill_manual(values = c("blue", "orange"),
#                     labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
#   theme_bw(base_size = 20) +
#   labs(y = expression("Solar radiation (MJ/m"^"2" * ")"),
#        x = element_blank()) +
#    guides(color = guide_legend(title = NULL),
#          fill = guide_legend(title = NULL)) +
#     theme(
#     panel.grid = element_blank(),
#     panel.background = element_rect(fill = "white"),
#     text = element_text(color = "black"),
#     axis.title = element_text(color = "black"),
#     axis.text = element_text(color = "black"),
#     strip.text = element_blank()
#   ) 
# 
# # Relative humidity plot
# rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean, 
#                                         color = site, group = site)) +
#   geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   scale_color_manual(values = c("blue", "orange"),
#                      labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
#   scale_fill_manual(values = c("blue", "orange"),
#                     labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
#   theme_bw(base_size = 20) +
#   labs(y = 'RH (%)', x= element_blank()) +
#    guides(color = guide_legend(title = NULL),
#          fill = guide_legend(title = NULL)) +
#     theme(
#     panel.grid = element_blank(),
#     panel.background = element_rect(fill = "white"),
#     text = element_text(color = "black"),
#     axis.title = element_text(color = "black"),
#     axis.text = element_text(color = "black"),
#     strip.text = element_blank()
#   ) 
# 
# # Print the plots
# t_moplot
# sol_moplot
# rh_moplot
# 
# 
# # t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean, 
# #                                       color = site, group = site)) +
# #   geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.2) +
# #   geom_line() +
# #   geom_point() +
# #   theme_bw(base_size=14) +
# #   labs(y = expression("Mean temperature " ( degree~C)),
# #        x = element_blank())
# # 
# # t_moplot
# # 
# # sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean, 
# #                                       color = site, group = site)) +
# #   geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.2) +
# #   geom_line() +
# #   geom_point() +
# #   theme_bw(base_size = 14) +
# #   labs(y = expression("Solar radiation (MJ/m"^"2" * ")"),
# #        x = element_blank()) 
# # 
# # sol_moplot
# # 
# # 
# # rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean, 
# #                                       color = site, group = site)) +
# #   geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.2) +
# #   geom_line() +
# #   geom_point() +
# #   theme_bw(base_size = 14) +
# #   labs(y = 'RH (%)', x= element_blank()) 
# # rh_moplot
# library(ggpubr)
# 
# quad <- ggarrange(p_cols, 
#           t_moplot,
#           sol_moplot,
#           rh_moplot,
#           common.legend = T,
#           legend = 'bottom',
#           ncol=2,
#           nrow=2) 
#           #nrow = 4)
#           #ncol= 0)
# quad
# 
# ggsave("./figures/quad.png", plot = quad, dpi = 600)
# 
# quad_plus <- ggarrange(flow_plot2,quad,
#                        nrow=2)
# quad_plus
# 
# ggsave("/Users/megansears/Documents/Repos/fourmile/figures/quad_plus_20250329.png", plot = quad_plus, dpi = 600)
# 
# 

@@

# HRU climate diff

@@

```{r}
#
# statn_catch <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_model/output/4milecreek38HRUs-statnall_#1/out/Catchment.csv',
#                         skip=3) %>%
#   slice(3:n()) %>%
#   dplyr::select(-starts_with('@'),
#                 -1) %>%
#   mutate(time = ymd(time)) %>%
#   mutate_if(is.character, as.numeric) %>%
#   mutate(source = 'statn')
#
# grid_catch <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_model/output/4milecreek38HRUs-gridall_#1/out/Catchment.csv',
#                        skip=3) %>%
#   slice(3:n()) %>%
#   dplyr::select(-starts_with('@'),
#                 -1) %>%
#   mutate(time = ymd(time)) %>%
#   mutate_if(is.character, as.numeric) %>%
#   mutate(source = 'grid')
#
# catchment_p <- bind_rows(statn_catch, grid_catch) %>%
#   dplyr::select(precip, time, source) %>%
#   pivot_wider(., names_from=source, values_from=precip)
#
# p_comp <- ggplot(catchment_p, aes(x=statn, y=grid)) + geom_point(size = 3) +
#   theme_bw(base_size=25) +
#   labs(x = 'Station precip (mm)',
#        y = 'Grid precip (mm)') +
#   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")
#
# p_comp
# ggsave("./figures/p_compare.png", plot = p_comp, dpi = 600)

@@

p_cols <- catchment_mo2 %>% 
  filter(month_name %in% c('Jun','Jul','Aug', 'Sep')) %>%
  ggplot(., aes(x = month_name, y = avg_p,
                                    color = site, fill = site)) +
  geom_col(position = 'dodge') +
  geom_errorbar(aes(ymin = 0, ymax = avg_p + sd_p), 
                position = position_dodge(width = 0.6), width = 0.2) +
  scale_color_manual(values = c("#0047AB", "orange"),
                     labels = c("Gridded", "Station")) +  # Rename color legend items
  scale_fill_manual(values = c("#0047AB", "orange"),
                    labels = c("Gridded", "Station")) +   # Rename fill legend items
  theme_bw(base_size = 20) +
  labs(y = 'Precipitation (mm)',
       x = element_blank()) +
  guides(color = guide_legend(title = NULL),
         fill = guide_legend(title = NULL)) +
    theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank()
  )

@@

ggsave("/Users/megansears/Documents/Repos/fourmile/figures/quad_20250329.png", plot = quad, dpi = 600)

quad_plus <- ggarrange(flow_plot2,quad,
                       nrow=2)
quad_plus

ggsave("/Users/megansears/Documents/Repos/fourmile/figures/quad_plus_20250329.png", plot = quad_plus, dpi = 600)
